Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils

### 2016-11-18 manuscript/response_to_reviewer_comments.tex

\medskip
{\it 1.8. Line 31-34: I cannot find discussion of this point in the rest of the text, and while important, this statement is relatively vague and there are no cited references. Since it underpins the rational for studying deep, unfrozen Arctic soils, it would be helpful to expand on this more in the manuscript. }

Referee \#2 raised this point as well (comment 2.6), and it's a good one. We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.  {\bf See new lines LINES.}

\medskip
{\it 1.9. Lines 48-60: I think that this section would benefit from an introduction of the interactions between the specific ecosystem (upland boreal forest) you are studying, and its interaction with soil chemistry, since vegetation type is influential in terms of soil carbon quality and quantity. Woody plant biomass tends to have a higher C:N ratio relative to herbaceous dominated systems, and this tends to result in lower quality resources for microbial communities. }

Vegetation and ecosystem type is a significant factor that is not well explored here, we agree. We have added some points about this in the introduction and discussion. {\bf See new lines LINES.}

@@

\medskip
{\it 3.6. 29: Not really sure how the comparison as to the relative controls of T and moisture are evaluated. }

This statement has been reworded to remove the comparison. {\bf See new lines LINES.}

\medskip
{\it 3.7. 50: see also updates in Hugelius et al. (2014)
63: Under some conditions (Olefeldt et al 2013): vague and confusing. Please clarify. }

Reviewer 2 also raised the issue (comment 2.4) of our incomplete citation of relevant literature. The Tarnocai reference has been replaced by one to Hugelius et al. (2014), and the Olefeldt sentence clarified. {\bf See new lines LINES.}

### 2016-11-18 manuscript/cpcrw_manuscript.Rmd
Final tweaks to new Table 1
Comment 1.9 - vegetation-soil interactions

The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such any such feedback remain highly uncertain {Schuur, 2015 #4319}. While northern soils hold enormous quantities {Tarnocai, 2009 #2231} of potentially mineralizable soil organic carbon (SOC), vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}. Such dynamics may however be disrupted by increased fire disturbance, particularly with more-intense fires {Johnstone, 2010 #2465}{Genet, 2013 #4412}. In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}{Karhu, 2014 #4124}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.

Temperature and moisture typically have strong and often interactive influences on soil GHG emissions. Laboratory incubations, field observations, and meta-analyses have documented increased fluxes of CO~2~, and under some conditions of CH~4~ {Olefeldt, 2013 #3456}, with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}. Greenhouse gas (GHG) responses to wetting and thawing dynamics are much less certain, with substantial variability between studies arising from differences in soil type, antecedent conditions, phase changes, experimental protocols, etc. {Kim, 2012 #2863}. The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}. Decadal warming and drying trends in Alaska {Bieniek, 2014 #4389} may counteract these effects, however.

@@

discuss - changes over time (see new reuslts)

>>>>

The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such any such feedback remain highly uncertain {Schuur, 2015 #4319}. While northern soils hold enormous quantities {Hugelius, 2014 #3978} of potentially mineralizable soil organic carbon (SOC), vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}. Vegetation type also influences SOC quality and quantity, with microbial communities {Högberg, 2007 #4482}, soil respiration {Raich, 2000 #596}, and SOC  all linked to aboveground factors such as woody versus nonwoody stems, deciduous versus evergreen canopies, and the presence of nitrogen-fixing plants.

These feedbacks between vegetation type, ground cover, permafrost, and SOC may be disrupted by increased fire disturbance, particularly after more-intense fire events {Johnstone, 2010 #2465}{Genet, 2013 #4412}. Even absent disturbance, the stability of SOC is highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}{Karhu, 2014 #4124}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.

Temperature and moisture typically have strong and often interactive influences on soil GHG emissions. Laboratory incubations, field observations, and meta-analyses have documented changing greenhouse gas (GHG) fluxes with rising temperature {Olefeldt, 2013 #3456}{Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}. GHG responses to wetting and thawing dynamics exhibit substantial variability between studies, probably due to differences in soil type, antecedent conditions, phase changes, experimental protocols, etc. {Kim, 2012 #2863}. The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}. Decadal warming and drying trends in Alaska {Bieniek, 2014 #4389} may counteract these effects, however.

@@

discuss - changes over time (see new reuslts)
discuss - upland mixed-canopy boreal forest here; careful about extrapolation
